---
title: "Exercise: Explore, Clean, and Wrangle Data"
author: "Webb"
format: html
editor: visual
---

## About the data
# Package ID: knb-lter-sbc.77.8 Cataloging System:https://pasta.edirepository.org.
# Data set title: SBC LTER: Reef: Abundance, size and fishing effort for California Spiny Lobster (Panulirus interruptus), ongoing since 2012.
# Data set creator:    - Santa Barbara Coastal LTER
# Data set creator:  Daniel C Reed & Robert J Miller -  

```{r}
inUrl1  <- "https://pasta.lternet.edu/package/data/eml/knb-lter-sbc/77/8/f32823fba432f58f66c06b589b7efac6" 
infile1 <- tempfile()
try(download.file(inUrl1,infile1,method="curl"))
if (is.na(file.size(infile1))) download.file(inUrl1,infile1,method="auto")
```



## Setup 
```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(tidyr)
```
# Read in data and explore data 

```{r}
# Read in data
lobster_abundance <- read_csv("https://portal.edirepository.org/nis/dataviewer?packageid=knb-lter-sbc.77.8&entityid=f32823fba432f58f66c06b589b7efac6")

lobster_traps <- read_csv("https://portal.edirepository.org/nis/dataviewer?packageid=knb-lter-sbc.77.8&entityid=66dd61c75bda17c23a3bce458c56ed84")

head(lobster_abundance)
```
```{r}
head(lobster_abundance)
```

## Exercise 1 Convert missing values using mutate() and na_if()

```{r}
lobster_abundance <- lobster_abundance %>% 
    mutate(SIZE_MM = na_if(SIZE_MM, -99999))
head(lobster_abundance)
```

## Exercise 2 filter() practice
```{r}
aque_70mm <- lobster_abundance %>% 
    filter(SITE == "AQUE" & SIZE_MM >= 70)
head(lobster_abundance)
```

## Exercise 3 Create a subset (vector) with the traps’ information at all sites where abundance data is not NA.*** REVIEW***
```{r}
## Create a vector with unique sites in lobster_abundance
abundance_sites <- unique(lobster_abundance$SITE)
head(abundance_sites)

## Filter sites in vector above
traps_subset <- lobster_traps %>% 
    filter(SITE %in% abundance_sites)
head(traps_subset)
```


## Exercise 4 Calculate totals by site and year
```{r}
total_abundance <- lobster_abundance %>% 
    group_by(SITE, YEAR) %>% 
    summarize(total_lobsters = sum(COUNT, na.rm = T))


total_traps <- traps_subset %>% 
    group_by(SITE, YEAR) %>% 
    summarize(total_traps = sum(TRAPS, na.rm = T))
```


## Exercise 5 Joining two data frames total_abundance and total_traps

```{r}
head(total_abundance)
head(total_traps)
```


## Exercise 6 ## Adding a new column
